Disc stuff

### 2012-10-31 paper.tex
some new figs model description section written up

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{graphicx}

@@

The model used to simulate the emission from a young protoplantary disc is a hybrid model created from hydrodynamical models of an infalling prestellar core and a gravitationally unstable protoplanetary disc. The prestellar core model is based upon the model of Keto \& Caselli (2010). This describes an infalling pre-stellar core, with densities temperatures and velocities given by the model of the collapse of a 10 solar mass Bonnor-Ebert sphere and provides a match to line profiles seen in the pre-stellar core L1544. Some slight modification due to the addition of Oxygen lines (Keto et al 2012 in prep) have been been added to this model resulting in a slight decrease in the peak infall velocity. This prestellar core model extends from 80$\,$au out to 10,000$\,$AU. The model is a 1D spherically symmetric model with inward motions (see figure \ref{l1544model}).

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/L1544model_used.png}
 \label{l1544model}
 \caption{envelope model showing temp and dust temp (red, green) in kelvin, log number density (blue) in cm$^{-3}$ and inward velocity$\times$10 (cyan) in m$\,$s$^{-1}$, adapted from Keto \& Caselli (2010)}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/rhoT.png}
 \label{discModel}
 \caption{The density and temperatures of the 0.4$\,\rm{M}_\odot$ gravitationally unstable disc model used in inside 80$\,$AU. The dense and hot regions are confinded to a few AU about the rotation plane.}
\end{figure*}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/velocity_rzSlice_thetaColourScale2.png}
 \label{discVel}
 \caption{The line of sight (colour scale) and plane of sky (arrows) velocities in the y=0 slice through the disc model.}
\end{figure*}

At radii less than 80$\,$au the physical structure is the given by the hydrodynamic model described in Ilee et al. (2011), based on the work of Boley(2007), and Boley(2009). This model describes a 0.39$\,\rm{M}_\odot$ self-gravitating disc featuring prominent spiral arms. H$_2$ number densities in the disc range from 10$^{10}$-10$^{19}\,\rm{m}^{-3}$, and temperatures range from 30-400 K. The dust and gas temperatures in the disc are assumed to be in equilibrium. The model is sampled over a regular grid of size 256$\times$256$\times$64 with spatial resolution of 0.5$\,$au in x, y and z. The majority of the mass lies in the mid-plane of the disc. For the models using a smoothed disc the same physical and chemical model is used but with the temperature, density and abundance averaged in $\phi$. The gas/dust mass ratio was assumed to be 1/100 throughout both sections model and the opacities were given by the model of dust grains with thick icy mantles and 10$^6\,$yr coagulation from Ossenkopf and Henning (1994)\newline

Chemical abundances in the disc were taken from Ilee et. al (2011) which followed the changes of chemical abundances of trace particles moving through the disc as it evolved. The abundances of 125 species related by 1334 reactions were followed through the time evolution of the disc. These abundances were interpolated onto a 51$^3$ grid covering the disc with cells of size 2.2$\times$2.2$\times$0.22$\,$au. Abundances in the envelope model were as follows:\newline

(tableify this?)\newline
HCO$^+$: as the H2O profile from (find out where L1544 H2O profile is from) scaled so that the maximum abundance is 10$^{-8}$\newline
HNO: constant abundance of 5$\times$10${-11}$\newline
HCS$^+$: 10$^{-11}$\newline
OCS: constant abundance of 10$^{-9}$\newline
H$_2$CO: 1.5$\times$10$^{-9}$ decreased by a factor of 40 at radii less than 8250au (Young et al 2004)\newline
CS:  3$\times$10$^{-9}$ decreased by a factor of 10,000 at radii less than 6700au (Tafalla, santiago-garcia and myers 2006)\newline
CO: as the H2O profile from before scaled so that the maximum abundance is CO maximum\newline

\section{description of radiative transfer program}
LIME (Brinch 2011), the radiative transfer program used, calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3d model. The method of selecting these points is given in the griding section. At each of these points the density of the main collision partner (in this case H$_2$), gas and dust temperatures, velocity, molecular abundances and turbulent velocity and taken from the model. These points are then smoothed by Lloyds algorithm (Lloyd 1982) in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is down these paths that photons are restricted to propagating (figures \ref{grid}). The levels of molecules in question are then  calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value.\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}
 \label{grid}
 \caption{An example of the points distribution chosen by lime for a model of a smooth protoplanetary disc. The nodes of the grid are the points at which the underlying physical model is sampled and the edges are teh paths down which photons can travel during the excitation-radiation emission stage of the computation.}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/lime_points_rz2.png}
 \label{points}
 \caption{the rz distribution of points selected to sample the model. The points in the dense region at 10$^9<\rm{z}<10^12$ and 7*10$^{11}<\rm{r}<10^{13}$ are sampling the disc model and the points at larger values are those foreced into the outer regions to sample the envelople.}
\end{figure}

@@

@@

Continuum emission from the model. Able to do this as the majority of the computing time on the model is taken up with calculating molecule population levels. This image has been created by averaging the output images of 100 unique runs. This averaging mitigates the granularity of output images from lime due to the nature of the gridding.

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/continuum.png}

 \caption{A composite image of 100 runs of continuum emission at 350$\,$GHz}
\end{figure}

Figures showing the RT results in a few molecules/transitions (CO, HCO$^+$, HCN, OCS, H2CO, NH3 --- maybe H2O,H2$^{18}$O [to try first]) (Note the moment 1 and 0 map were created by integrating between -12.5 to -0.5 km$\,$s$^{-1}$ and +0.5 to +12.5 km$\,$s$^{-1}$ to avoid being dominated by the contribution from the envelope, this can be seen in the some PV diagrams as the strong absorption feature at all positions around zero velocity, mom1maps are shown with a cutoff of 1/1000 of the peak emission/absorption value)\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_contSub.png}

 \caption{CO 3-2 Continuum subtracted mom0}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_mom1.png}

 \caption{CO 3-2 mom1map}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_PV_centre.png}

 \caption{CO 3-2 PV through centre}
\end{figure}

These synthetic images of the CO 3-2 line at 345.8$\,$GHz. As the upper level for this line is at 33$\,$K above the ground state this transition should be excited through out disc. CO line will be thick so this is only mapping the outer regions of the disc. Spiral structure can be seen.\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageOCS_28-27_30deg_contSub.png}

 \caption{OCS 28-27 Continuum subtracted mom0}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageOCS_28-27_30deg_PV_centre.png}

@@

The OCS lines in alma band 7 (22-21 through 30-29) with upper energy levels between 161 and 271 K above the ground state provide a way to trace hot shocked gas in spiral arms without resolving structure.\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imagesmoothedOCS_28-27_30deg_contSub.png}

 \caption{SmoothedOCS 28-27 Continuum subtracted mom0}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imagesmoothedOCS_28-27_30deg_PV_centre.png}

 \caption{SmoothedOCS 28-27 PV through centre}
\end{figure}

In comparison to the model with spiral arms the absorbtion and emission from this smoothed model is far weaker.\newline

@@

line ratios of many lined spieces can help us calculate temperatures of the emitting regions.

(figure of temperature as a function of line ratio for 2 sets of formaldehyde lines and map of derived temperature, talk to P about this)\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_1-0_30deg_contSub.png}

 \caption{HCO$^+$ 1-0 Continuum subtracted mom0}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_1-0_30deg_mom1.png}

 \caption{HCO$^+$ 1-0 mom1map}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_1-0_30deg_PV_centre.png}

 \caption{HCO$^+$ 1-0 PV through centre}
\end{figure}

some molecules such as HCO$^+$ trace only the inner disc (ref ilee 2011) and so can be used to look at the extended velocity and physical structure.\newline

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_mom1.png}

 \caption{C18O 3-2 15 deg mom1map}
\end{figure}

@@

fdagfdabfdsbfdbds fdbfd  fdfbfvdfd g rbfd fdbfdbc \newline fdagfdabfdsbfdbds fdbfd  fdfbfvdfd g rbfd fdbfdbc \newline fdagfdabfdsbfdbds fdbfd  fdfbfvdfd g rbfd fdbfdbc \newline fdagfdabfdsbfdbds fdbfd  fdfbfvdfd g rbfd fdbfdbc \newline fdagfdabfdsbfdbds fdbfd  fdfbfvdfd g rbfd fdbfdbc \newline fdagfdabfdsbfdbds fdbfd  fdfbfvdfd g rbfd fdbfdbc \newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_mom1.png}

 \caption{C18O 3-2 30 deg mom1map}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_mom1.png}

 \caption{C18O 3-2 45 deg mom1map}
\end{figure}

@@

nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline nfsdankjdsankj vdjsk nvjdks nvkds jkkd jkvd jkvf djk fj\newline

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_mom1.png}

 \caption{C18O 3-2 60 deg mom1map}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_mom1.png}

 \caption{C18O 3-2 75 deg mom1map}
\end{figure}

@@

Figure showing different transitions of the same molecule (e.g. CO(1-0), ...(7-6) , OCS,  H2CO ) for same inclination but in different disks (Boley et al. and the smooth disk)\newline

@@

- current status (Cycle 1) - Figure
OCS + C18O + H2CO + HNO/CS/
- final status - Figure

\section{Conclusions}

note that all species show up in absorption and some show a little bit of emission around the edges of the disc
OCS is good for showing up spiral structure without being able to resolve it.
envelope only contaminates the central plus-minus 500 m/s or so

@@

I thank Professor N. Kameswara Rao for some helpful suggestions,
Dr H. C. Bhatt for a critical reading of the original version of the
paper and an anonymous referee for very useful comments that improved
the presentation of the paper.

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{graphicx}
\usepackage{float}

@@

%-describe the envelope structure (pre-stellar core)
The model used to simulate the emission from a young protoplantary disc is a hybrid model created from hydrodynamical models of an infalling prestellar core and a gravitationally unstable protoplanetary disc. The prestellar core model is based upon the model of Keto \& Caselli (2010). This describes an infalling pre-stellar core, with densities temperatures and velocities given by the model of the collapse of a 10 solar mass Bonnor-Ebert sphere and provides a match to line profiles seen in the pre-stellar core L1544. Some slight modification due to the addition of Oxygen lines (Keto et al 2012 in prep) have been been added to this model resulting in a slight decrease in the peak infall velocity. This prestellar core model extends from 80$\,$au out to 10,000$\,$AU. The model is a 1D spherically symmetric model with inward motions (see figure \ref{l1544model}).\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/L1544model_used.png}
 \label{L1544}
 \caption{Model of the prestellar core L1544 used as the envelope in the hybrid model. Showing temperature and dust temperature (red, green) in kelvin, log number density (blue) in cm$^{-3}$ and inward velocity$\times$100 (cyan) in m$\,$s$^{-1}$. Adapted from Keto \& Caselli (2010)}
\end{figure}

\begin{figure*}
 \includegraphics[width=168mm]{Figures/model/rhoT2.png}
 \label{rhoT} 
 \caption{Left: A 3D plot of log density (m$^3$) showing the spiral structure in the xy plane and scale height of the disc. Right: The 3D temperature strucure of the disc; regions less than 40 degrees are not shown in 3D, demonstrating the narrow central region containing hot material.}
\end{figure*}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/rotational_velocities.png}
 \label{velocity}
 \caption{Azimuthaly avereaged rotational velocity in the disc midplane with best fit curve}
\end{figure}

%-describe the physical structure of the disk (density and temperature) -- one figure showing the physical structure and the kinematics (2-panels figure)
At radii less than 80$\,$au the physical structure is given by the hydrodynamic model described in Ilee et al. (2011), based on the work of Boley(2007), and Boley(2009). This model describes a 0.39$\,\rm{M}_\odot$ self-gravitating disc featuring prominent spiral arms. H$_2$ number densities in the disc range from 10$^{10}$-10$^{19}\,\rm{m}^{-3}$, and temperatures range from 30-400 K (figure \ref{rhoT}). The dust and gas temperatures in the disc are assumed to be in equilibrium. The model is sampled over a regular grid of size 256$\times$256$\times$64 with spatial resolution of 0.5$\,$au in x, y and z. The majority of the mass lies in the mid-plane of the disc. For the models using a smoothed disc the same physical and chemical model is used but with the temperature, density and abundance averaged in $\phi$. The gas/dust mass ratio was assumed to be 1/100 throughout both sections model and the opacities were given by the model of dust grains with thick icy mantles and 10$^6\,$yr coagulation from Ossenkopf and Henning (1994)\newline

%-describe the chemistry  - refer to Ilee et al. which have been taken as input to the rad transfer (RT) code
Chemical abundances in the disc were taken from Ilee et. al (2011) which followed the changes of chemical abundances of trace particles moving through the disc as it evolved. The abundances of 125 species related by 1334 reactions were followed through the time evolution of the disc. These abundances were interpolated onto a 51$^3$ grid covering the disc with cells of size 2.2$\times$2.2$\times$0.22$\,$au. Abundances in the envelope model were as follows:\newline

(tableify this?)\newline
HCO$^+$: as the H2O profile from Caselli (2012, submitted) scaled so that the maximum abundance is 10$^{-8}$\newline
HNO: constant abundance of 5$\times$10${-11}$\newline
HCS$^+$: 10$^{-11}$\newline
OCS: constant abundance of 10$^{-9}$\newline
H$_2$CO: 1.5$\times$10$^{-9}$ decreased by a factor of 40 at radii less than 8250au (Young et al 2004)\newline
CS:  3$\times$10$^{-9}$ decreased by a factor of 10,000 at radii less than 6700au (Tafalla, santiago-garcia and myers 2006)\newline
CO: as the H2O profile from Caselli (2012, submitted) scaled so that the maximum abundance is CO maximum\newline

%-describe the RT used (LIME) 
LIME (Brinch 2011), the radiative transfer program used, calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3d model. The method of selecting these points is given in the griding section. At each of these points the density of the main collision partner (in this case H$_2$), gas and dust temperatures, velocity, molecular abundances and turbulent velocity and taken from the model. These points are then smoothed by Lloyds algorithm (Lloyd 1982) in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is down these paths that photons are restricted to propagating (figures \ref{grid}). The levels of molecules in question are then  calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value.\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/Lime_grid3.png}
 \label{grid}
 \caption{A plot of the points selected by the gridding process and the paths down which photons can propagate overlaid on a smoothed density model. The points are more concentrated at small radii and in the densest regions.}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/model/lime_points_rz2.png}
 \label{points}
 \caption{A plot of the positions of the points selected to sample the model}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/average_comparison.png}
 \label{averages}
 \caption{A plot showing the difference between a single LIME continuum image at 300$\,$GHz and the average of ten such images from different runs of the same model in LIME}
\end{figure}

@@

%Figures showing the RT results in a few molecules/transitions (CO, HCO$^+$, HCN, OCS, H2CO, NH3 --- maybe H2O,H2$^{18}$O [to try first])
The results of these simulations for the molecules OCS, HCO$^+$, H$_2$CO and isotopologues of CO are shown here, other molecules simulated but not shown include HCN, HNC, HNO, HCS$^+$ and  CS.
For the purpose of simulating observations the model was placed at roughly the distance of nearby low-mass star forming regions (100$\,$pc) and is inclined 30$^\circ$ to edge on. From these observations integrated intensity, intensity weighted velocity and position velocity diagrams through the centre of the model were created.
The simulations done are focused upon frequencies within ALMA band 7, selected to give the best trade off between resolution and sensitivity for early ALMA science.
(Note the moment 1 and 0 maps were created by integrating between -12.5 to -0.5 km$\,$s$^{-1}$ and +0.5 to +12.5 km$\,$s$^{-1}$ to avoid being dominated by the contribution from the envelope, this can be seen in some PV diagrams as the strong absorption feature at all positions around zero velocity, moment 1 maps are shown with a cutoff of 1/1000 of the peak emission/absorption value)\newline

\begin{figure}[H]
 \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_contSub.png}

 \caption{CO J=3-2 Continuum subtracted mom0}
\end{figure}

\begin{figure}[H]
 \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_mom1.png}

 \caption{CO J=3-2 moment 1 map}
\end{figure}

\begin{figure}[H]
 \includegraphics[width=84mm]{Figures/sim/imageCO_3-2_30deg_PV_centre.png}

 \caption{CO J=3-2 PV through y=0''}
\end{figure}

Figures 6,7 and 8 show synthetic images of the CO J=3-2 line at 345.8$\,$GHz. As the upper level for this line is at 33$\,$K above the ground state this transition should be excited through out disc. given the large and uniform nature of the abundance of CO in the disc this line will be optically thick and this is only mapping the outer regions of the disc. In common with most transitions simulated, the line is seen almost exclusively in absorption of the disc continuum. The region of emission in the centre is where there is no continuum background to be absorbed.  Some indications of spiral structure can be seen but are indistinct. ***not sure how usefull these are, might be worth cutting them?***\newline

\begin{figure}[H]
 \includegraphics[width=84mm]{Figures/sim/imageOCS_28-27_30deg_composite_contSub.png}
 \label{OCS_mom0}
 \caption{OCS 28-27 Continuum subtracted mom0}
\end{figure}

\begin{figure}[H]
 \includegraphics[width=84mm]{Figures/sim/imageOCS_28-27_30deg_composite_PV_centre.png}

@@

The OCS lines in alma band 7 (22-21 through 30-29) with upper energy levels between 161 and 271 K above the ground state provide a way to trace hot shocked gas in spiral arms without resolving structure. Figures \ref{OCS_mom0} \& \ref{smoothOCS_mom0} show the integrated intensity of the OCS J=28-27 line in the model described previously and one where the disc section of the model is an axis-symmetric model created by taking the azimuthal average of density, temperature and molecular abundance. The order of magnitude difference in the line intensity provides a method to detect hot dense gas without being able to reolve the spiral structure spatially.\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imagesmoothedOCS_28-27_30deg_contSub.png}
 \label{smoothOCS_mom0}
 \caption{SmoothedOCS 28-27 Continuum subtracted mom0}
\end{figure}

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imagesmoothedOCS_28-27_30deg_PV_centre.png}
%
% \caption{SmoothedOCS 28-27 PV through centre}
%\end{figure}

@@

More complex molecules such as H$_2$CO with many closely spaced spectral lines can be used to gain an estimate on the temperature of a region.

(figure of temperature as a function of line ratio for 2 sets of formaldehyde lines and map of derived temperature, talk to P about this. This doesnt seem doable with the lines around 1mm for alma)\newline

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_3-2_30deg_contSub.png}

 \caption{HCO$^+$ 3-2 Continuum subtracted mom0}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_3-2_30deg_mom1.png}

 \caption{HCO$^+$ 3-2 mom1map}
\end{figure}

\begin{figure}
 \includegraphics[width=84mm]{Figures/sim/imageHCOp_3-2_30deg_PV_centre_2.png}

 \caption{HCO$^+$ 3-2 PV through centre with rotation curve from figure \ref{velocity}}
\end{figure}

Some molecules such as HCO$^+$ trace only the outer regions of the disc (ref ilee 2011) and so can be used to look at the extended velocity and physical structure. In these colder less dense regions we see the molecular lines in emission rather than absorption \newline
\newline \newline

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_15deg_mom1.png}
%
% \caption{C18O 3-2 15 deg mom1map}
%\end{figure}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_30deg_mom1.png}
%
% \caption{C18O 3-2 30 deg mom1map}
%\end{figure}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_45deg_mom1.png}
%
% \caption{C18O 3-2 45 deg mom1map}
%\end{figure}

@@

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_60deg_mom1.png}
%
% \caption{C18O 3-2 60 deg mom1map}
%\end{figure}

@@

%\begin{figure}
% \includegraphics[width=84mm]{Figures/sim/imageC18O_3-2_75deg_mom1.png}
%
% \caption{C18O 3-2 75 deg mom1map}
%\end{figure}

@@

In all the simulations except HCO$^+$ we see molecular lines in absorption throughout the majority of the disc and in some cases a small amount of emission towards the centre, where the inclination of the disc allows us the view the low density but moderate temparture region inside the inner edge of the spiral structures.\newline

@@

%- current status (Cycle 1) - Figure
%OCS + C18O + H2CO + HNO/CS/
%- final status - Figure

casa continuum, OCS and HCO+ figs here

\section{Discussion}

In this paper we have performed radiative transfer simulations of a hybrid model comprising a 0.4$\, M_\odot$ self gravitating disc with radius 64$\,AU$ showing spiral density waves, surrounded by an envelope simulated as a collapsing 10$\,M_\odot$ BE-sphere. CASA simulations show that at a distance of 100$\,pc$ both spatial resolution of the spirals in such a disc and extraction of kinematic and temperature from molecular lines are possible in ALMA band 7. Our simulations show that many  molecular spieces are prodominantly seen in absorbtion towards the centre of  self-gravitating protoplanetary discs.\newline
The quiesient nature of the envelope around such discs only obscures them within $\pm\,0.5 km\,s^{-1}$.\newline
Hot, shocked gas can be detected in such a disc without resolving it by observing OCS transitions between 26-27 to 30-29.\newline
Mention that dust and gas are assumed to be in temperature equilibrium if they arent and dust is significantly cooler then transitions may not show up in absorption. atm spirals shock heat the mid-plane layer and the colder more difusse gas absorbs the continuum from it (need to ask aaron whether this is the case, the narrow midplane is much hotter than the rest but should check that this is due to shock heating, though with there being no accreation or iradiation i dont see what else it could be)\newline
note that the disc mass is on the high end for early discs\newline
rotation curves can be gatherered from these observations even with spiral structure\newline

@@

stuff here
